---
title: "MTF - getting started, testing out data import/export etc."
output:
  html_document:
    df_print: paged
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
require(haven)
```

I can use the `haven` package (part of the tidyverse) to read SAS files
```{r import_from_SAS, message = FALSE, warning = FALSE}
df1 = read_sas(data_file = "data/12th-grade-2016-2018/y2018_1.sas7bdat")
knitr::kable(df1[1:5,1:8])
```

Or just read a CSV exported from SAS.
*Note: I'm getting a lot of problems here, but I suspect it's from my original SAS work, rather than the importing step*
```{r import_from_CSV, message = FALSE, warning = FALSE}
df2 = read_csv(file = "data/ALLYEARS_81012.csv")
knitr::kable(df2[1:5,1:8])
```

# Goal #1: Reproduce a smaller version of [Table 1 from the MTF 2018 report](http://www.monitoringthefuture.org//pubs/monographs/mtf-overview2018.pdf#page=59&zoom=100,0,0)

Specifically: lifetime prevalence of use of a handful of different substances among all grades combined, for 2016-2018 (highlighted below).

## Step 1: get core data from all 3 years and grades into 1 table

I'll keep the cigarette, vaping, and alcohol indicators for "ever used" (lifetime prevalence) as well as "last 30 days": 
```{r twelve_core_3_years, warning = FALSE, message = FALSE}

# Question: what do I do if the thing I'm looking for isn't in that year's form?
twelve_core = function(year) {
  path = paste("data/12th-grade-2016-2018/y", year, "_1.sas7bdat", sep = "")
  
  basic_set = read_sas(data_file = path) %>% 
    select(.,
           id = RESPONDENT_ID,
           year = V1,
           sex = V2150,
           cig_ever = V2101,
           cig_month = V2102,
           alc_lifetime = V2104,
           alc_month = V2106,
           alc_drunk_lifetime = V2020,
           weight = ARCHIVE_WT,
           ) %>% 
    mutate(.,
           grade = 12,
           sex = recode_factor(sex, "1" = "Male", "2" = "Female", "-9" = NA_character_),
           cig_ever = recode_factor(cig_ever, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           cig_month = recode_factor(cig_month, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           alc_lifetime = recode_factor(alc_lifetime, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           alc_month = recode_factor(alc_month, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           alc_drunk_lifetime = recode_factor(alc_drunk_lifetime, "1" = "No", "-9" = NA_character_, .default = "Yes")
           )
  
  if (year >= 2017) { # the year they added in more nuanced vaping questions
    vaping_data = read_sas(data_file = path) %>% 
     select(.,
           id = RESPONDENT_ID,
           vape_ever = V2566,
           vape_nic_lifetime = V2567,
           vape_nic_month = V2569,
           vape_mj_lifetime = V2570,
           vape_mj_month = V2572,
           vape_flav_lifetime = V2573,
           vape_flav_month = V2575
          ) %>% 
      mutate(.,
           vape_ever = recode_factor(vape_ever, "0" = "No", "1" = "Yes", "-9" = NA_character_),
           vape_nic_lifetime = recode_factor(vape_nic_lifetime, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           vape_nic_month = recode_factor(vape_nic_month, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           vape_mj_lifetime = recode_factor(vape_mj_lifetime, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           vape_mj_month = recode_factor(vape_mj_month, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           vape_flav_lifetime = recode_factor(vape_flav_lifetime, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           vape_flav_month = recode_factor(vape_flav_month, "1" = "No", "-9" = NA_character_, .default = "Yes")
           ) %>% 
      mutate(., vape_any_month = as.factor(if_else(
                 (vape_nic_month == "Yes" | vape_mj_month == "Yes" | vape_flav_month == "Yes"), "Yes", "No")
                )
              )
  } else {
    vaping_data = read_sas(data_file = path) %>% 
     select(.,
           id = RESPONDENT_ID,
           vape_ever = V2551,
           vape_any_month = V2552
          ) %>% 
      mutate(.,
           vape_ever = recode_factor(vape_ever, "1" = "No", "-9" = NA_character_, .default = "Yes"),
           vape_nic_lifetime = NA_character_,
           vape_nic_month = NA_character_,
           vape_mj_lifetime = NA_character_,
           vape_mj_month = NA_character_,
           vape_flav_lifetime = NA_character_,
           vape_flav_month = NA_character_,
           vape_any_month = recode_factor(vape_any_month, "1" = "No", "-9" = NA_character_, .default = "Yes")
           )
  }
  output = full_join(basic_set, vaping_data, by = "id")
  
  output
}


# combine these together into full dataset - later add in other years
combined = tibble()
for (year in 2016:2018) {
  combined = bind_rows(combined, twelve_core(year))
}

knitr::kable(head(combined))
summary(combined[, c(2, 3, 4, 5, 6, 17, 18)])
```

```{r, echo = FALSE, eval = FALSE}
# require(table1)
```



# Goal #2: Reproduce trend graphs from the [NIDA for Teens interactive chart](https://teens.drugabuse.gov/teachers/stats-and-trends-teen-drug-use)

Specifically, I'd like to make [this view](https://public.tableau.com/shared/JF2B545PM?:display_count=y&:origin=viz_share_link&:embed=y), showing 8th, 10th, and 12th grade trends in past-30-day alcohol use and vaping, from 2016-2018.


# Goal #3: Use information from across multiple forms

???

# Questions:
* What should I do about the weights?
* Should I include missing data in the denominator for prevalences?
* For a question like *V2566: BY18 34230 EVER VAPE (p. 56)*, what does 70% missing mean? were some large amount of participants not asked? 